---
title: "HPFC Model Details"
author: "Alejandro Abraham"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HPFC Model Details}
  %\VignetteEngine{knitr::rmarkdown}
---

# Model Sample

This vignette demonstrates how to set up the required directories, load the relevant packages, and retrieve both spot and forward data for gas and power models.

It provides high level model wrapping flow.

## Setup

- Load the necessary packages
- Everything is conducted and determined in the `params.json` file. 

```{r setup, echo = TRUE}
box::use(
    data.table[...],
    magrittr[...],
    echarts4r[...],
    reactable[...],
    # HPFC[...],
    eikonapir[...]
)

devtools::load_all()

```

## Options

#### `**forecast_source**`
It will determine the price level source. It can be:
- `FWD`: commonly used to forecasting
- `DAM`: for backtesting purposes mainly.

#### `**data_source**`
- **Market**: it requires the Reuters API key
- **Manual**: parameter `MANUAL` and it requires the correct .xlsx files in the directory `sim_name` directory `/01_raw`

#### `**model_source**`
- **TRAIN**: it will train with the selected `data_source` a new set of models. 
- **LOCAL**: load the models from the `sim_name` directory `/02_output`.

## Load Inputs and Retrieve data

```{r, inputs, echo = TRUE}

LST_PARAMS = jsonlite::fromJSON(file.path('params.json'))

list_inputs = load_inputs(params = LST_PARAMS)
```


### Prepare Gas

```{r, prep-gas, echo = TRUE}
if(list_inputs$LST_PARAMS$model_source == 'TRAIN') {
    ENV_MODELS_GAS = prepare_gas(list_inputs = list_inputs)
}
```

### Prepare Power

```{r, prep-pwr, echo = TRUE}
if(list_inputs$LST_PARAMS$model_source == 'TRAIN') {
    ENV_MODELS_PWR = prepare_pwr(list_inputs = list_inputs)
}
```

## Train

### Train Gas - Long term

```{r, train-gas-lt, echo = TRUE}
if(list_inputs$LST_PARAMS$model_source == 'TRAIN') {

    ENV_MODELS_GAS$dt_lt_param_gasdep = 
      train_lt_gas(
        gas_data = ENV_MODELS_GAS$dt_lt_param_gasdep,
        ric_gas = unique(ENV_MODELS_GAS$dt_gas$RIC)
      )
    
    if(list_inputs$LST_PARAMS$archive != 'NO') {
        saveRDS(ENV_MODELS_GAS, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'ENV_MODELS_GAS.rds'))
        saveRDS(ENV_MODELS_GAS$dt_lt_param_gasdep, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'lt_gas_model.rds'))
        saveRDS(ENV_MODELS_GAS$dt_gas_dd_filt, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'saved_history_gas.rds'))
    }
    
}

if(list_inputs$LST_PARAMS$model_source == 'LOCAL') {
    if ("LST_DIRS" %in% names(list_inputs)) {
        ENV_MODELS_GAS = readRDS(file.path(list_inputs$LST_DIRS$dir_data_output, 'ENV_MODELS_GAS.rds'))
    }
    if ("LST_DIRS_archive" %in% names(list_inputs)) {
        ENV_MODELS_GAS = readRDS(file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'ENV_MODELS_GAS.rds'))
    }    
}

```

### Train Power - Long term

```{r, train-pwr-lt, echo = TRUE}
if(list_inputs$LST_PARAMS$model_source == 'TRAIN') {
    ENV_MODELS_PWR$dt_lt_param_pwr = 
      train_lt_pwr(
        pwr_data = ENV_MODELS_PWR$dt_lt_param_pwr,
        ric_pwr = unique(ENV_MODELS_PWR$dt_pwr$RIC),
        pwr_holidays = ENV_MODELS_PWR$calendar_holidays_pwr,
        gas_history = ENV_MODELS_PWR$gas_history
      )
}
    
```

### Train Power - Short term

```{r, train-pwr-st, echo = TRUE}
if(list_inputs$LST_PARAMS$model_source == 'TRAIN') {
    ENV_MODELS_PWR$lst_hr_param_pwr = 
      train_st_pwr(
        pwr_data = ENV_MODELS_PWR$dt_hr_param_pwr,
        gas_history = ENV_MODELS_PWR$gas_history
      )
    
    if(list_inputs$LST_PARAMS$archive != 'NO') {
        saveRDS(ENV_MODELS_PWR, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'ENV_MODELS_PWR.rds'))
        saveRDS(ENV_MODELS_PWR$dt_lt_param_pwr, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'lt_pwr_model.rds'))
        saveRDS(ENV_MODELS_PWR$dt_pwr_filt_dd, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'saved_history_gas.rds'))
        saveRDS(ENV_MODELS_PWR$lst_hr_param_pwr, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'st_pwr_model.rds'))
    }
    
}

if(list_inputs$LST_PARAMS$model_source == 'LOCAL') {
    if ("LST_DIRS" %in% names(list_inputs)) {
        ENV_MODELS_PWR = readRDS(file.path(list_inputs$LST_DIRS$dir_data_output, 'ENV_MODELS_PWR.rds'))
    }
    if ("LST_DIRS_archive" %in% names(list_inputs)) {
        ENV_MODELS_PWR = readRDS(file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'ENV_MODELS_PWR.rds'))
    }    
}
    
```


## Forecast 

### Preparation
```{r, pred-prep, echo = TRUE}
LST_FOR = list(
    model_lt_gas = copy(ENV_MODELS_GAS$dt_lt_param_gasdep),
    model_lt_pwr = copy(ENV_MODELS_PWR$dt_lt_param_pwr),
    model_st_pwr = copy(ENV_MODELS_PWR$lst_hr_param_pwr),
    dt_fwds = copy(list_inputs$ENV_FWD$dt_fwds),
    saved_history_gas = copy(list_inputs$ENV_SPOT$history_gas),
    saved_history_pwr = copy(list_inputs$ENV_SPOT$history_pwr),
    ric_spot_gas = list_inputs$ENV_SPOT$spot_gas_RIC,
    ric_fwd_gas = unique(HPFC::spot_GAS_products_full[products_GAS %in% c(LST_PARAMS$selected_gas_code, LST_PARAMS$dependent_gas_code)]$products_GAS_code),
    ric_spot_pwr = list_inputs$ENV_SPOT$spot_pwr_RIC,
    ric_fwd_pwr = unique(HPFC::spot_PWR_products_full[countries %in% LST_PARAMS$selected_pwr_code]$products_PWR_code),
    calendar_forecast = list_inputs$ENV_CODES$calendar_future,
    start_date = LST_PARAMS$forecast_start,
    end_date = LST_PARAMS$forecast_end,
    last_date = list_inputs$ENV_CODES$last_date
) 

```

### Predict Gas 

```{r, pred-gas, echo = TRUE}
ENV_FOR_GAS = forecast_gas(input_forecast = LST_FOR)

if(list_inputs$LST_PARAMS$archive != 'NO') {
    saveRDS(ENV_FOR_GAS, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'forecast_gas.rds'))
}
```

### Predict Power 

```{r, pred-pwr, echo = TRUE}
ENV_FOR_PWR = forecast_pwr(input_forecast = LST_FOR, gas_forecast = ENV_FOR_GAS)

if(list_inputs$LST_PARAMS$archive != 'NO') {
    saveRDS(ENV_FOR_PWR, file.path(list_inputs$LST_DIRS_archive$dir_data_output, 'forecast_pwr.rds'))
}
```

## Visualization 

### Prepare data for Backtesting

```{r, viz, echo = TRUE}
dt_pwr_for = ENV_FOR_PWR[, .(date, hour, forecast = final_forecast, RIC, season, peak, value_gas, value_bl = spot_forward_month_BL)]
dt_pwr_obs = HPFC::dt_spot_pwr[year(date) %in% unique(year(dt_pwr_for$date)) & RIC == unique(LST_FOR$ric_spot_pwr)][, .(date, hour, spot = value, RIC)]
dt_pwr = merge(dt_pwr_for, dt_pwr_obs, by = c('date', 'hour', 'RIC'), all = TRUE)

setcolorder(dt_pwr, c('date', 'hour', 'season', 'peak', 'RIC', 'spot', 'forecast', 'value_bl', 'value_gas'))
setorder(dt_pwr, date, hour)

```

```{r, table data-backtesting}
reactable(dt_pwr)
```

### Daily Curve - Spot, Forecast & Commodities

```{r, viz-predictedobserved}
dt_pwr_lg = melt(dt_pwr, id.vars = c('date', 'hour', 'season', 'peak', 'RIC'), variable.name = 'type', value.name = 'value')
dt_pwr_lg[, .(value = round(mean(value))), by = .(date, type)] %>% 
  group_by(type) %>% 
  e_charts(date) %>% 
  e_line(value, smooth = TRUE, symbol='none') %>% 
  e_legend(show = TRUE, orient = 'horizontal') %>% 
  e_title(paste(unique(dt_pwr_lg$RIC), "Mean Daily Curve", 'Whole period ')) %>% 
  e_tooltip(trigger = "axis") %>% 
  e_toolbox_feature(feature = "saveAsImage", title = "Save as image") %>% 
  e_datazoom(start = 0) %>% 
    e_theme('westeros')

```


```{r}

```

